Reproduction Survey Analysis

### 2022-09-28 procedure/code/203_results.Rmd
markdown of analysis
I split the analysis into an Rmarkdown document (203_results.Rmd), keeping pre-processing in 202_analysis.R  (which could be renamed now)

## R Markdown

Abstract...

@@

```{r sub-fields-by-methods}
# crosstab sub-fields and methods
ftable(data = survey_resp, Q4 ~ Q3)
```

```{r familiarity-by-subfield-and-method}
# crosstab familiarity with sub-fields and methods
ftable(data = survey_resp, Q5 ~ Q3 + Q4)
```

```{r familiarity-by-subfield}

# calculate familiarity with reproducible practices
survey_resp <- survey_resp %>% mutate(familiar = as.numeric(Q7a) + 
                                        as.numeric(Q7b) + as.numeric(Q7c) +
                                        as.numeric(Q7d) + as.numeric(Q7e))

# summarize and ANOVA familiarity with practices by subfield
group_by(survey_resp, Q3) %>%
  summarise(
    count = n(),
    mean = mean(familiar, na.rm = TRUE),
    sd = sd(familiar, na.rm = TRUE)
  )
```

@@

#### Summarize reproducible research practices ####
# there must be a prettier option for a crosstabulation with marginal sums!

summary(survey_resp$Q7a)
summary(survey_resp$Q7a_1)
ftable(data = survey_resp, Q7a_1 ~ Q7a)
summary(survey_resp$Q7a_3)
ftable(data = survey_resp, Q7a_3 ~ Q7a)
# FAMILIARITY with open source software outstrips USE of open source software
# and of documenting the computational environment
# low percentage of researchers "always" share information about computational environment

summary(survey_resp$Q7b)
summary(survey_resp$Q7b_1)
ftable(data = survey_resp, Q7b_1 ~ Q7b)
# FAMILIARITY with notebooks also outstrips their USE 
# less than 17% of researchers always use notebooks (lab, field, or computational)
# less than 40% use them most of the time

summary(survey_resp$Q7c)
summary(survey_resp$Q7c_1)
ftable(data = survey_resp, Q7c_1 ~ Q7c)
summary(survey_resp$Q7c_2)
ftable(data = survey_resp, Q7c_2 ~ Q7c)
summary(survey_resp$Q7c_3)
ftable(data = survey_resp, Q7c_3 ~ Q7c)
# most researchers are familiar with sharing/archiving data, but still only 16%
# report always doing so
# use of DOIs is inconsistent and use of metadata is rare

summary(survey_resp$Q7d)
summary(survey_resp$Q7d_1)
ftable(data = survey_resp, Q7d_1 ~ Q7d)
summary(survey_resp$Q7d_2)
ftable(data = survey_resp, Q7d_2 ~ Q7d)
# most researchers are at least somewhat familiar with sharing code or scripts
# but less than 40% actually do so all or most of the time
# and only 14% use version control software some or most of the time

summary(survey_resp$Q7e)
summary(survey_resp$Q7e_1)
ftable(data = survey_resp, Q7e_1 ~ Q7e)
# familiarity with pre-analysis plan registration is rare:
# 48% of respondents have no familiarity
# only 6 respondents reported using pre-analysis registrations all or most of the time

# In Sum, most geographers are at least somewhat familiar with open source software,
# data sharing, and code sharing.
# there are large gaps in implementation across all practices
# some practices are practically unknown in geography, including metadata, 
# version control software, and pre-analysis plan registration.

##### Summarize Beliefs #####

survey_resp %>% select(starts_with("Q8")) %>% summary()

# calculate familiarity with reproducible practices
survey_resp <- survey_resp %>% mutate(belief = as.numeric(Q8_1) + 
                                        (4 - as.numeric(Q8_2)) + 
                                        as.numeric(Q8_3) +
                                        as.numeric(Q8_4) + 
                                        as.numeric(Q8_5) +
                                        (4 - as.numeric(Q8_6))
                                      )
# a majority of geographers believe a failed reproduction does not disprove
# the original study, and geographers are split on whether it detracts from
# the original study's validity
# the most agreement is for students attempting reproductions as part of training

survey_resp %>% select(starts_with("Q8")) %>% lapply(as.numeric) %>% lapply(summary, na.rm=TRUE)
# we can see that in aggregate, geographers do not tend to strongly agree or disagree
# with beliefs about reproducibility
# most in favor of students conducting replications, and reproducibility contributing
# to credibility, and even some degree of compatibility with epistemology in the field
# however geographers are, on average, neutral about loss of validity for failed reproductions
# or loss of trust for studies without available data. 
# we disagree with a failed reproduction falsifying the original study.

# summarize and ANOVA beliefs with practices by subfield
group_by(survey_resp, Q3) %>%
  summarise(
    count = n(),
    mean = mean(belief, na.rm = TRUE),
    sd = sd(belief, na.rm = TRUE)
  )

aov(belief ~ Q3, data=survey_resp) %>% summary()
# again, significant differences in aggregate beliefs about the role of 
# reproducibility in subfields, with physical holding strongest affirmative
# beliefs, then methods, nature/society, and human geography

>>>>

## Reproducibility Survey Results

First, load the pre-processed data.

@@

### What research methods does each sub-field use?

The methods and statistics subfield primarily uses quantitative methods, and secondarily uses mixed methods.
The physical geography subfield uses quantitative and mixed methods, likely including qualitative researcher
interpretations as "mixed methods".
The human geography and nature/society subfields use qualitative, quantitative and mixed methods, with the majority preferring qualitative or mixed methods.

```{r sub-fields-by-methods}
# crosstab sub-fields and methods
addmargins(table(survey_resp$Q4, survey_resp$Q3))
```

### How familiar is each subfield with the term "reproducibility"?

Physical and Methods subfields are more familiar with the
term "reproducibility" than the Nature/Society or Human subfields.

```{r familiarity-by-subfield}
addmargins(table(survey_resp$Q5, survey_resp$Q3))
```

### How familiar is each method with the term "reproducibility"?

Quantitative and Mixed Methods researchers are more familiar with the
term "reproducibility" than qualitative researchers.

```{r familiarity-by-methods}
addmargins(table(survey_resp$Q5, survey_resp$Q4))
```

### How do geographers define reproducibility?

The following code creates a results table, `definitions.csv`, organizing
definitions by researcher's self-reported familiarity with the term reproducibility
and their preferred methodology (quant, qual, or mixed).

We can see a trend where the more familiar a researcher is with the term "reproducibility", 
the more likely they are to report a definition consistent with the NASEM.
The less familiar a researcher is, the more likely they are to report a definition
consistent with "replication", whereby different data may be used and results are 
expected to be similar, but not identical, to the original results.

```{r define-reproducibility}
# how did folks define "reproducibility" ?
survey_resp %>% select(Q3, Q4, Q5, Q6) %>%
  filter(Q5 != "Not at all") %>%
  arrange(Q5, Q4) %>%
  write.csv(here("results", "tables", "definitions.csv"))
```

### How familiar is each subfield with reproducible research practices?

The following analysis aggregates familiarity with several different research 
practices into a single indicator, including:
1. use of open source software,
2. research notebooks (lab, field, or computational),
3. data sharing,
4. code sharing, and 
5. pre-analysis research protocol registration.

Familiarity is measured from 1 (not at all) to 4 (to a great extent) for each
research practice, resulting in possible scores from 5 (not familiar with any practice)
to 20 (greatly familiar with every practice).

The methods and physical geography subfields report the most familiarity with
reproducible research practices, followed by nature/society and lastly human geography.

```{r familiarity-by-subfield}
group_by(survey_resp, Q3) %>%
  summarise(
    count = n(),
    mean = round(mean(familiar, na.rm = TRUE),2),
    sd = round(sd(familiar, na.rm = TRUE),2)
  )
```

There are significant differences between subfields with regards to familiarity with reproducible research practices.

@@

## Does familiarity with practices translate into applying them to research?

### Do researchers familiar with open source software use it in research?

Although a majority of researchers are familiar with open source software, they
less frequently use the software for research.

```{r open-source-use}
addmargins(table(survey_resp$Q7a, survey_resp$Q7a_1))
```

### What do researchers use open source software for?

```{r open-source-use}

## variables not ready for analysis

```

### Do researchers familiar with open source software share the computational environment?

Although a majority of researchers are familiar with open source software, they
less frequently share the details of the computational environment used in their
research.

```{r environment-share}
addmargins(table(survey_resp$Q7a, survey_resp$Q7a_3))
```

### Do researchers familiar with research notbooks use them in their research?

FAMILIARITY with notebooks exceeds their use in research. 
Less than 17% of researchers always use notebooks (lab, field, or computational)
Less than 40% use them most of the time.
If researchers are at best somewhat familiar with notebooks, then they only
use them some of the time.

```{r notebook-use}
addmargins(table(survey_resp$Q7b, survey_resp$Q7b_1))
```

### Do researchers familiar with sharing data do so in their research?

Very few researchers always share their data, and familiarity with sharing
data exceeds the frequency with which researchers do so.

```{r data-share}
addmargins(table(survey_resp$Q7c, survey_resp$Q7c_1))
```

Even for those sharing data most of the time, they do not necessarily use
digital object identifiers (DOIs) to persistently link to the data.

```{r doi}
addmargins(table(survey_resp$Q7c, survey_resp$Q7c_2))
```

And an even larger number of researchers rarely if ever document spatial metadata
for the data that they are sharing, nullifying its usefulness for others.
Just 16% of researchers report always documenting spatial metadata.

```{r metadata}
addmargins(table(survey_resp$Q7c, survey_resp$Q7c_3))
```

### Do researchers familiar with sharing code do so in their research?

Most researchers are at least somewhat familiar with sharing code or scripts,
but less than 40% actually do so all or most of the time

```{r share-code}
addmargins(table(survey_resp$Q7d, survey_resp$Q7d_1))
```

Only 14% use version control software some or most of the time

```{r version-control}
addmargins(table(survey_resp$Q7d, survey_resp$Q7d_2))
```

### Do researchers familiar with pre-analysis planning register pre-analysis plans for their research?

Only 52% of respondents have any familiarity with pre-analysis plans, and of those
just 6 researchers register plans all or or most of the time.

```{r pre-analysis}
addmargins(table(survey_resp$Q7e, survey_resp$Q7e_1))
```

In Sum, most geographers are at least somewhat familiar with open source software,
data sharing, and code sharing.
there are large gaps in implementation across all practices
some practices are practically unknown in geography, including metadata, 
version control software, and pre-analysis plan registration.

# Beliefs about Reproducibility

```{r beliefs-summary}
survey_resp %>% select(starts_with("Q8")) %>% summary()
```

A majority of geographers believe a failed reproduction does not disprove the original study,
and geographers are split on whether it detracts from the original study's validity.
The most agreement is for students attempting reproductions as part of training

```{r beliefs-stats}
survey_resp %>% select(starts_with("Q8")) %>% lapply(as.numeric) %>% lapply(summary, na.rm=TRUE)
```

we can see that in aggregate, geographers do not tend to strongly agree or disagree
with beliefs about reproducibility
most in favor of students conducting replications, and reproducibility contributing
to credibility, and even some degree of compatibility with epistemology in the field
however geographers are, on average, neutral about loss of validity for failed reproductions
or loss of trust for studies without available data. 
we disagree with a failed reproduction falsifying the original study.

# Do subfields vary in their aggregate beliefs about reproducibility?

There are differences in aggregate beliefs about the role of 
reproducibility between the four subfields, with physical holding strongest affirmative
beliefs, then methods, nature/society, and human geography

```{r beliefs-stats-by-subfield}
group_by(survey_resp, Q3) %>%
  summarise(
    count = n(),
    mean = mean(belief, na.rm = TRUE),
    sd = sd(belief, na.rm = TRUE)
  )
```

The differences are statistically significant.

```{r beliefs-anova}
aov(belief ~ Q3, data=survey_resp) %>% summary()
```
